Analysis PPM-S

### 2020-07-31 docs/paper.Rmd
update-julio
correción de tablas, títulos y figuras.

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use,  empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy. Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    template: null
    toc: false 
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

```{r include=FALSE}
# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
# clean #in the yml
```

@@

```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
pacman::p_load(knitr,kableExtra, dplyr)
```

@@

<div style="text-align: justify">

# (PART\*) Introducción {-}

@@

* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.

* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.

* H3. Meritocratic preferences are a latent variable based on a normative value of effort and talent.

* H4. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.

@@

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A.Perception Effort" =perc_effort,
                          "B.Perception Talent" = perc_talent,
                          "C.Perception rich parents" = perc_wpart,
                         "D.Perception contacts" = perc_netw,
                          "E.Preferences Effort" = pref_effort,
                          "F.Preferences Talent" =  pref_talent,
                         "G.Preferences rich parents" =  pref_wpart ,
                          "H.Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2, font_size = 10,position = "center",latex_options = c("hold_position"));sumdat04
```

El gráfico XX muestra ...
A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.

@@

El Grafico \@ref(fig:corpoly) muestra ....
Se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces. Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

@@

Figure X shows the results of the estimation for the four-factor model:

> Julio: Esta sección debe ser revisada:

- CFA general
- Fit x modelo

[Aquí SEM plot con cargas y ajuste global del modelo]

Los modelos confirmatorios fueron estimados en primer lugar para cada uno de los tres administration-sets. Los resultados se presentan en la Tabla \@ref(tab:tab-loads)

Como se puede observar en la Tabla XX, todos los modelos independiente de los órdenes obtuvieron un ajuste adecuado. con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3. El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).  Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968,RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).

@@

> Julio: ¿Esto se debe mantener?

```{r include=FALSE}
cap <- "Items according to dimension."
# if (knitr::is_html_output()) {te <- table_nums("te",cap)} else {te <- cap}
```

|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
|:---------:|:-----------------------:|:---------------------------:|:---------------:|
| **n**     |           1769          |             1769            |       1769      |
| **CFI**   |          0.595          |            0.988            |      0.994      |
| **TLI**   |          0.433          |            0.976            |      0.985      |
| **RMSEA** |          0.226          |            0.047            |      0.036      |
| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
| **p**     |           .000          |             .000            |       .000      |
| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
Table:`r cap`

@@

# Appendix {-}

```{r include=FALSE}
cap <- "Representativeness of the sample."
# if (knitr::is_html_output()) {aax <- table_nums("aax",cap)} else {aax <- cap}
```

|                     | Sample  |   CEP  |
|---------------------|:-------:|:------:|
| __Gender__          |         |        |
| Men                 | 49,82%  | 50,52% |
| Women               | 50.18%  | 49,47% |
| __Age__             |         |        |
| 18 - 24             | 18,55%  | 18,17% |
| 25 - 34             | 18,86%  | 17,48% |
| 35 - 44             | 19.09%  | 19,98% |
| 45 - 54             | 17,96%  | 19,23% |
| 55 - or more        | 25,54%  | 25.11% |
| __Education__       |         |        |
| Primary  or less    | 2,93%   | 15,88% |
| Hig school          | 43,23%  | 37,04% |
| Non university      | 32,63%  | 28,93% |
| university or more  | 21,21%  | 18,13% |
Table:`r cap`

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use,  empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy. Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false 
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

```{r eval=FALSE, include=FALSE}
# for render in pdf run rmarkdown::render_site("docs/paper.Rmd", output_format = "all")
# clean #in the yml
rmarkdown::render("docs/paper.Rmd", output_format = "bookdown::pdf_document2")
```

@@

```{r message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
pacman::p_load(knitr, kableExtra, dplyr, lavaan)
```

@@

# Introducción

@@

* $H_1$. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.

* $H_2$. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.

* $H_3$. Meritocratic preferences are a latent variable based on a normative value of effort and talent.

* $H_4$. Non-meritocratic preferences are a latent variable based on the normative value of the use of personal contacts and having rich parents.

@@

```{r desc01, echo=FALSE}
cap <- "Descriptive statistics of the scale."
load(file = "../input/data/proc/dat04.RData")
dat04b<- dat04 %>% rename("A. Perception Effort" =perc_effort,
                          "B. Perception Talent" = perc_talent,
                          "C. Perception rich parents" = perc_wpart,
                          "D. Perception contacts" = perc_netw,
                          "E. Preferences Effort" = pref_effort,
                          "F. Preferences Talent" =  pref_talent,
                          "G. Preferences rich parents" =  pref_wpart ,
                          "H. Preferences contacts" = pref_netw  )
sum04b  <- psych::describe(dat04b[,-c(1)], fast = T) %>% select(mean,sd,min,max) %>% data.frame() %>% round(digits = 2)
table04b<- sum04b %>% rename("Mean"=mean, "SD"=sd, "Min"=min,"Max"=max)
sumdat04<- kable(table04b,format =table_format, booktabs= T, escape = FALSE,align = "l",linesep = "", caption = cap
) %>% kable_styling(full_width = table_format2, font_size = 10,position = "center",latex_options = c("hold_position"));sumdat04
```

A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.

@@

En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces. Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

@@

In \@ref(fig:meas02) the results of the estimation for the four-factor model with complete sample. This model have a good fit (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88. Furthermore, considering the correlation between the latent variables, it is observed that the most related variables are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.

```{r meas02, echo=FALSE, fig.cap = "Theorical model", fig.align='center'}
knitr::include_graphics('../output/images/meas02.png')
```

Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado. con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3. El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).  Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).

```{r tab03, echo=FALSE,eval=TRUE}
load(file = "../output/tables/fit4_o.RData")
load(file = "../output/tables/fit3_o.RData")
load(file = "../output/tables/fit2_o.RData")
load(file = "../output/tables/fit1_o.RData")

sum_fit<- bind_rows(fitmeasures(fit4_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
                    fitmeasures(fit1_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
                    fitmeasures(fit2_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")],
                    fitmeasures(fit3_o)[c("chisq.scaled","df","cfi.scaled","rmsea.scaled")])
sum_fit$mod <- c("Model 1","Model 2","Model 3","Model 4")
sum_fit$est <- c("DWLS","DWLS","DWLS","DWLS")
sum_fit$nobs <- c(nobs(fit4_o),nobs(fit1_o),nobs(fit2_o),nobs(fit3_o))
sum_fit <- select(sum_fit,mod,nobs,est,everything())
colnames <- c("Model","$N$","Estimator","$\\chi^2$","df","CFI","RMSEA")

sumtable02<- kable(sum_fit,digits = 3,format = table_format,row.names = F,booktabs=T, align = c("lcccccc"),
                   caption = "Summary fit indices according group",
                   col.names = colnames,
                   escape = FALSE) %>%
  kable_styling(full_width = F,
                font_size = 10)  %>%
      footnote(number = c("Model 1: Complete sample",
                          "Model 2: fixed order by percepction/preference",
                          "Model 3:  fixed order by topic (i.e: effort)",
                          "Model 4: Randomized order"));sumtable02
```

@@

|           | Contrast Model 1 Factor | theoretical model 4 Factors | Model with M.I. |
|:---------:|:-----------------------:|:---------------------------:|:---------------:|
| **n**     |           1769          |             1769            |       1769      |
| **CFI**   |          0.595          |            0.988            |      0.994      |
| **TLI**   |          0.433          |            0.976            |      0.985      |
| **RMSEA** |          0.226          |            0.047            |      0.036      |
| **$\chi2$**    |         1830.839        |            68.661           |      40.250     |
| **p**     |           .000          |             .000            |       .000      |
| **$\chi2$/df** |          65.38          |             4.90            |       3.32      |
Table: `r as.character(paste("(\\#tab:items-dim)", "Items according to dimension."))`.

@@

# (APPENDIX) Appendix {-}

`r if (knitr::is_latex_output()){ '\\appendix'}`

`r if (knitr::is_latex_output()){ '\\section{Appendix}'}`

|                     | Sample  |   CEP  |
|---------------------|:-------:|:------:|
| __Gender__          |         |        |
| Men                 | 49,82%  | 50,52% |
| Women               | 50.18%  | 49,47% |
| __Age__             |         |        |
| 18 - 24             | 18,55%  | 18,17% |
| 25 - 34             | 18,86%  | 17,48% |
| 35 - 44             | 19.09%  | 19,98% |
| 45 - 54             | 17,96%  | 19,23% |
| 55 - or more        | 25,54%  | 25.11% |
| __Education__       |         |        |
| Primary  or less    | 2,93%   | 15,88% |
| Hig school          | 43,23%  | 37,04% |
| Non university      | 32,63%  | 28,93% |
| university or more  | 21,21%  | 18,13% |
Table: `r as.character(paste("(\\#tab:rep-samp)", "Representativeness of the sample"))`.

### 2020-07-31 production/prod_analysis-cfa.rmd
update-julio
correción de tablas, títulos y figuras.

semPaths(fit4_o , 
whatLabels = "std",
what="col",
label.cex=1,
edge.label.cex=1,
residuals=F, 
edge.color = "black",
style = "lisrel",
fade=F, 
nCharNodes = 0,
 curvePivot = FALSE,
         curve = 2,
rotation = 4,
layout = "tree3",
         cardinal = "lat cov",  
legend.cex = 0.6,
         label.cex =1,
label.font = 6,
         edge.label.font= 1,
         asize = 3,
edge.width =1,
         sizeMan=5,
         sizeLat = 10, width = 12,
         height= 15, nodeNames = nodeNames, intercepts = F, reorder = F, thresholds = F, nodeLabels = c("A","B","C","D","E","F","G","H","1","2","3","4"), filetype ="png", filename = "output/images/meas02") #arrow size)

@@

dat_merit_2<- dat_merit %>% sjmisc::rec(rec="rev") %>% select("a.Effort"=a.Effort_r,"a.Talent"=a.Talent_r, "a.Rich Parents"="a.Rich Parents_r", "a.Contacts"=a.Contacts_r, "b.Effort"=b.Effort_r,"b.Talent"=b.Talent_r,"b.Rich Parents"="b.Rich Parents_r", "b.Contacts"=b.Contacts_r)

plotlikert<-plot_likert(dat_merit_2, c(1,1,1,1,2,2,2,2),
            groups.titles = c("Perceptions","Preferences"),
            geom.colors ="PuBu", 
            geom.size = 0.8,
            axis.labels=c("Effort","Talent","Rich Parents","Contacts"),
            catcount = 4,
            cat.neutral = 3, 
            grid.range  =  c ( 1.2 , 1.4 ),
            values  =  "sum.outside", reverse.colors = T)
plotlikert

ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 25,height = 15,dpi = "retina",units = "cm")

>>>>

semPaths(
  fit4_o ,
  whatLabels = "std",
  # what = "col",
  label.cex = 1,
  edge.label.cex = 0.7,
  residuals = F,
  optimizeLatRes =T,
  edge.color = "black",
  style = "lisrel",
  fade = F,
  nCharNodes = 0,
  curvePivot = FALSE,
  curve = 3,
  rotation = 4,
  layout = "tree2",
  cardinal = "lat cov",
  legend.cex = 0.6,
  label.cex = 1,
  label.font = 6,
  edge.label.font = 7,
  asize = 3,
  edge.width = 1,
  sizeMan = 5,
  sizeLat = 10,
  width = 20,
  height = 15,
  nodeNames = nodeNames,
  intercepts = F,
  reorder = F,
  thresholds = F,
  fixedStyle =1,
  nodeLabels = c("A", "B", "C", "D", "E", "F", "G", "H", "1", "2", "3", "4"),
  filetype = "png",
  filename = "output/images/meas02"
) #arrow size)

@@

dat_merit_2<- dat_merit %>% sjmisc::rec(rec="rev") %>% 
  select(
    "a.Effort" = a.Effort_r,
    "a.Talent" = a.Talent_r,
    "a.Rich Parents" = "a.Rich Parents_r",
    "a.Contacts" = a.Contacts_r,
    "b.Effort" = b.Effort_r,
    "b.Talent" = b.Talent_r,
    "b.Rich Parents" = "b.Rich Parents_r",
    "b.Contacts" = b.Contacts_r)

plotlikert<-plot_likert(dat_merit_2, 
                        c(1, 1, 1, 1, 2, 2, 2, 2),
                        groups.titles = c("Perceptions", "Preferences"),
                        geom.colors = "PuBu",
                        geom.size = 0.8,
                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                        catcount = 4,
                        cat.neutral = 3,
                        grid.range  =  c (1.2 , 1.4),
                        values  =  "sum.outside", 
                        reverse.colors = T,
                        reverse.scale = F)

plotlikert

ggsave(plotlikert,filename = "output/images/plotlikert.png",device = "png",width = 30,height = 15,dpi = "retina",units = "cm")
```
